AggregateCounts

### 2021-06-16 R/AggregateCounts.Rmd
Added figures from previous paper and got the readTiff function to work

#now plot it
ggplot()+
    geom_jitter(data = newHistogramSummary.df,aes(x=as.factor(LogBin),y=NormalizedFrequency,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Frequency",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))  
#####
#now only aggregates 5 microns or larger
#####
aggsOnlyHist.df<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% filter(!LogBin == 5) %>% group_by(Species,Condition,TimePoint, Replicate)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>% group_by(Species,Condition,TimePoint, Replicate, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

@@

#now plot it
#normalized frequency
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedFrequency,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Frequency",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet timepoint, color species
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet species, color timepoint
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass,  color =TimePoint, ),size=3, height = 0, width = .2)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~Species)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet species and Timepoint, color Replicate
ggplot()+
    geom_line(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass, color = Replicate, group = Replicate ),size=1)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    guides(color = FALSE)+
    facet_grid(TimePoint~Species)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    scale_colour_grey()+
    theme(plot.title = element_text(hjust = 0.5))

#####
#Planktonic vs aggregates
#####
#Add agregation value to the dataframe
newHistogram.df<-newHistogram.df %>%  mutate(Aggregation = case_when(LogBin == 5 ~ "Planktonic",LogBin>5 ~ "Aggregate")) %>%
  group_by(Species,Condition,TimePoint, Replicate, Aggregation)%>% 
  mutate(AgPlBiomass = sum(Size), AgPlFreq = n()) %>% ungroup() %>%
  mutate(AgPlNormFrequency = AgPlFreq/TotalFrequency*100, AgPlNormBiomass = AgPlBiomass/TotalBiomass*100)
#Sumarize in dataframe
plankAggHist<-newHistogram.df %>% group_by(Species,Condition,TimePoint, Replicate, Aggregation,AgPlNormFrequency,AgPlNormBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkAggFreq =sum(AgPlNormFrequency), checkAggBiomass = sum(AgPlNormBiomass))
#plot Frequency
ggplot()+
    geom_jitter(data = plankAggHist,aes(x=as.factor(Aggregation),y=AgPlNormFrequency,  color =Species, ),size=3, height = 0, width = .2)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Frequency",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
ggplot()+
    geom_jitter(data = plankAggHist,aes(x=as.factor(Aggregation),y=AgPlNormBiomass,  color =Species, ),size=3, height = 0, width = .2)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))

#####
```

```{r fancy histogram of aggregates}
#Create histograms
##this takes a long time, so consider running them in parts
allDataGrouped<-aggregateList %>% group_by(Species, TimePoint, Condition)#first do the correct grouping
#allDataGrouped<-aggregateList %>% group_by(Species, TimePoint, Condition, Sample)# this is to include sample on the distribution or keep it out
allDataList<- group_split(allDataGrouped)#create list of dataframes
allDataHistograms.5<-map2(allDataList, 5,TotalFancyHistogram)#apply histogram function to all items in the list
allDataHistograms.5df<- bind_rows(allDataHistograms.5)#bring them together

@@

```{r biomass per slice}
#this was taken from "Microscopy data clean.Rmd"
GetResults<-function(sample,){#points to folder with files. Folder is hardcoded, unfortunately
  #folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Binary Images/', sample)
  folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Binary Images/wtd1-02')
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  blueChannel<-fileList[grep('B_',fileList)]
  for(slice in 1:length(filesList)/3){
    counts<-sum(as.array(readImage(fileList[1], convert = TRUE)))
    growthDf<-rbind(growthDf,c(counts, slice,species))
    growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
} }
#this was taken from the PropOc package
SlicesTo3D<-function(filesList,side){
  image1<-readTiff(filesList[1])#read first image to get its dimensions
  image3D <- array(0, c(image1@size[1], image1@size[2], length(filesList)))#preallocate array for image given using image dimensions
  for(i in 1:length(filesList)/3){#keep in mind that it is not zero-index
    #print(paste("loading ch1...", i, "/", length(filesList), filesList[i]))
    image3D[,,i]<-as.array(readImage(fileList[1], convert = TRUE))
  }
  return(image3D)
}

>>>>

#Plot Normalized Frequency
ggplot()+
    geom_jitter(data = newHistogramSummary.df,aes(x=as.factor(LogBin),y=NormalizedFrequency,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Frequency (%)",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
#Plot Normalized Biomass
ggplot()+
    geom_jitter(data = newHistogramSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Biomass (%)",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))

#Plot normalized Biomass per sample
ggplot()+
    geom_line(data = newHistogramSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass, color = Replicate, group = Replicate ),size=1)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    guides(color = FALSE)+
    facet_grid(TimePoint~Species)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    scale_colour_grey()+
    theme(plot.title = element_text(hjust = 0.5)) 
#####
#now only aggregates 5 microns or larger
#####
aggsOnlyHist.df<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% filter(!LogBin == 5) %>% group_by(Species,Condition,TimePoint, Replicate)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>% group_by(Species,Condition,TimePoint, Replicate, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

@@

#now plot it
#normalized frequency
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedFrequency,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Frequency (%)",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet timepoint, color species
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Biomass (%)",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet species, color timepoint
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass,  color =TimePoint, ),size=3, height = 0, width = .2)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~Species)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet species and Timepoint, color Replicate
ggplot()+
    geom_line(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass, color = Replicate, group = Replicate ),size=1)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    guides(color = FALSE)+
    facet_grid(TimePoint~Species)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    scale_colour_grey()+
    theme(plot.title = element_text(hjust = 0.5))

#####
#Planktonic vs aggregates
#####
#Add agregation value to the dataframe
newHistogram.df<-newHistogram.df %>%  mutate(Aggregation = case_when(LogBin == 5 ~ "Planktonic",LogBin>5 ~ "Aggregate")) %>%
  group_by(Species,Condition,TimePoint, Replicate, Aggregation)%>% 
  mutate(AgPlBiomass = sum(Size), AgPlFreq = n()) %>% ungroup() %>%
  mutate(AgPlNormFrequency = AgPlFreq/TotalFrequency*100, AgPlNormBiomass = AgPlBiomass/TotalBiomass*100)
#Collapse to only values per replicate
plankAggHist<-newHistogram.df %>% group_by(Species,Condition,TimePoint, Replicate, Aggregation,AgPlNormFrequency,AgPlNormBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkAggFreq =sum(AgPlNormFrequency), checkAggBiomass = sum(AgPlNormBiomass))
#plot Frequency
ggplot()+
    geom_jitter(data = plankAggHist,aes(x=as.factor(Aggregation),y=AgPlNormFrequency,  color =Species, ),size=3, height = 0, width = .2)+ 
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Frequency (%)",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Plot Biomass
ggplot()+
    geom_jitter(data = plankAggHist,aes(x=as.factor(Aggregation),y=AgPlNormBiomass,  color =Species, ),size=3, height = 0, width = .2)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Biomass (%)",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#####
```

```{r Figures from Pasa SCFM2 Paper}
#Biomass contribution of Planktonic vs Aggregates
#####
#Collapse dataframe to values per condition
plankAggHistSummary<-plankAggHist %>% group_by(Species,TimePoint, Condition,Aggregation) %>%
  summarise(MeanFreq = mean(AgPlNormFrequency), SEFreq = std.error(AgPlNormFrequency), MeanBiomass = mean(AgPlNormBiomass),
            SEBiomass = std.error(AgPlNormBiomass))
#Plot using the code from the paper
ggplot()+
  geom_col(data = plankAggHistSummary,aes(x=Aggregation,y=MeanBiomass, fill = TimePoint),
           size=3,position = position_dodge( width = .9))+
  geom_jitter(data = plankAggHist,aes(x=Aggregation,y=AgPlNormBiomass, group = TimePoint),shape = 1,size=2,position = position_dodge(width = .9),show.legend = FALSE)+
  geom_errorbar(data = plankAggHistSummary,aes(x=Aggregation,ymin=(MeanBiomass-SEBiomass), ymax = (MeanBiomass+SEBiomass), group = TimePoint),width = 0, size=.5,position = position_dodge(width = .9))+
  scale_y_continuous(limits = c(0,103),expand = expand_scale(mult = c(0, 0)))+
  facet_grid(.~Species)+
  scale_fill_grey()+
  labs(title = '', x = "", y = "Biomass contribution (%)")+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5),strip.text = element_text(size = 9, face = "italic",margin = margin(.1,0,.1,0, "cm")),
        strip.background = element_rect(color="black", fill="white", size=.5, linetype="solid"),
        axis.title.x = element_text(margin = margin(0,0,0,0),size=13),axis.text.x = element_text(margin = margin(1,0,0,0),size=13),
        axis.title.y = element_text(margin = margin(0,4,0,0),size=13),axis.text.y = element_text(margin = margin(0,0,2,0),size=11),
        legend.title = element_text(size=13),legend.text = element_text(size = 11),
          legend.position = c(0.79, 0.60))
#####
#Normalized Frequency and Biomass contribution of Aggregate bins
#####
#Collapse to data per condition
AggsOnlySummary<-aggsOnlyHistSummary.df %>% group_by(Species,TimePoint, Condition,LogBin) %>%
  summarise(MeanFreq = mean(NormalizedFrequency), SEFreq = std.error(NormalizedFrequency), MeanBiomass = mean(NormalizedBiomass),
            SEBiomass = std.error(NormalizedBiomass))
#Plot Frequency using code from the paper
ggplot()+
    geom_col(data = AggsOnlySummary, aes(x=as.factor(LogBin),y=MeanFreq,  fill =TimePoint ),size=4,position = position_dodge( width = .9))+
    geom_jitter(data = aggsOnlyHistSummary.df, aes(x=as.factor(LogBin),y=NormalizedFrequency, group = TimePoint),shape = 1,size=2,position = position_dodge( width = .9))+
    geom_errorbar(data = AggsOnlySummary,aes(x=as.factor(LogBin),ymin=(MeanFreq-SEFreq), ymax = (MeanFreq+SEFreq), group = TimePoint),width = 0, size= .5,position = position_dodge( width = .9))+
    facet_grid(.~Species)+
    labs(fill = "",y = "Aggregates per bin (% of total)",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    #guides(color = FALSE)+
    scale_y_continuous(limits = c(0,70), expand = expand_scale(mult = c(0, 0)))+
    scale_fill_grey()+
    theme_half_open()+
    theme(plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 9, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=.5, linetype="solid"),
          plot.margin = unit(c(0,.5,0,.5), "cm"),
          axis.title.x = element_text(margin = margin(0,0,0,0),size=13),axis.text.x = element_text(margin = margin(1,0,0,0),size=11),
          axis.title.y = element_text(margin = margin(0,4,0,0),size=13),axis.text.y = element_text(margin = margin(0,2,2,0),size=11),
          legend.title = element_text(size=13),legend.text = element_text(size = 11),
          legend.position = c(0.79, 0.60))
#Plot Normalized Biomass
ggplot()+
    geom_col(data = AggsOnlySummary, aes(x=as.factor(LogBin),y=MeanBiomass,  fill =TimePoint ),size=4,position = position_dodge( width = .9))+
    #geom_jitter(data = aggsOnlyHistSummary.df, aes(x=as.factor(LogBin),y=NormalizedBiomass, group = TimePoint),shape = 1,size=2,position = position_dodge( width = .9))+
    geom_errorbar(data = AggsOnlySummary,aes(x=as.factor(LogBin),ymin=(MeanBiomass-SEBiomass), ymax = (MeanBiomass+SEBiomass), group = TimePoint),width = 0, size= .5,position = position_dodge( width = .9))+
    facet_grid(.~Species)+
    labs(fill = "",y = "Aggregates per bin (% of total)",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    #guides(color = FALSE)+
    scale_y_continuous(limits = c(0,100), expand = expand_scale(mult = c(0, 0)))+
    scale_fill_grey()+
    theme_half_open()+
    theme(plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 9, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=.5, linetype="solid"),
          plot.margin = unit(c(0,.5,0,.5), "cm"),
          axis.title.x = element_text(margin = margin(0,0,0,0),size=13),axis.text.x = element_text(margin = margin(1,0,0,0),size=11),
          axis.title.y = element_text(margin = margin(0,4,0,0),size=13),axis.text.y = element_text(margin = margin(0,2,2,0),size=11),
          legend.title = element_text(size=13),legend.text = element_text(size = 11),
          legend.position = c(0.79, 0.60))
#####
```{r fancy histogram of aggregates}
#Create histograms
##this takes a long time, so consider running them in parts
allDataGrouped<-aggregateList %>% group_by(Species, TimePoint, Condition)#first do the correct grouping
#allDataGrouped<-aggregateList %>% group_by(Species, TimePoint, Condition, Sample)# this is to include sample on the distribution or keep it out
allDataList<- group_split(allDataGrouped)#create list of dataframes
allDataHistograms.5<-map2(allDataList, 5,TotalFancyHistogram)#apply histogram function to all items in the list
allDataHistograms.5df<- bind_rows(allDataHistograms.5)#bring them together

@@

```{r biomass per slice}
#this was taken from "Microscopy data clean.Rmd"
GetResults<-function(){
  folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Binary Images/wtd1-02')
  fileList<-list.files(path = folder, full.names = TRUE)
  blueChannel<-fileList[grep('B_',fileList)]
  redChannel<-fileList[grep('R_',fileList)]
  greenChannel<-fileList[grep('G_',fileList)]
  Results<-rbind(blueChannel,redChannel,greenChannel)
  Results<-Results %>% mutate(Sample = sample)
IterateTiff<-function(fileList, species)
  growth<-matrix(ncol=3, nrow=length(fileList))
  for(slice in 1:(length(fileList))){
    counts<-sum(as.array(readImage(fileList[slice], convert = TRUE)))
    growth[slice,]<-c(counts, slice, species)
  }
  growthDf<-data.fram(growth)
return(growthDf) 
  }

testGetResults<-GetResults()
testGetResults.df<- data.frame(testGetResults)
#this was taken from the PropOc package
SlicesTo3D<-function(filesList,side){
  image1<-readTiff(filesList[1])#read first image to get its dimensions
  image3D <- array(0, c(image1@size[1], image1@size[2], length(filesList)))#preallocate array for image given using image dimensions
  for(i in 1:length(filesList)/3){#keep in mind that it is not zero-index
    #print(paste("loading ch1...", i, "/", length(filesList), filesList[i]))
    image3D[,,i]<-as.array(readImage(fileList[1], convert = TRUE))
  }
  return(image3D)
}
